# Introduction {.unnumbered}


## Required packages

The following packages may be used during the course; it is assumed
that you know how to install packages, and have permission to do
so on your computer. 

### R version

We require at least R 4.1.0 (as we'll be using the native pipe `|>`), but it is recommended to always work with the latest released  version of R.

### CRAN packages:

```{r eval=FALSE}
install.packages(c(
"sf",
"stars",
"terra",
"s2",
"gstat",
"spatstat",
"spdep",
"spatialreg",
"mgcv"
))
```

### non-CRAN packages:

`starsdata` and `spDataLarge`:
```{r eval=FALSE}
options(timeout = 3600) # 1 hr instead of 1 min
install.packages("spDataLarge", repos = "https://nowosad.github.io/drat/", 
				 type = "source") # 23 Mb
install.packages("starsdata", repos = "http://cran.uni-muenster.de/pebesma/", 
				 type = "source") # 1 Gb!
```

### Course schedule

Schedule Thu:

* 9:00 - 10:00 lecture part 1a
* 10:00 - 10:15 coffee break
* 10:15 - 11:00 lecture part 1b
* 11:00 - 12:00 exercise 1 [answers](part1_ex.html)
* 12:00 - 13:00 lunch break
* 13:00 - 14:00 lecure part 2a
* 14:00 - 15:00 exercises 2a
* 14:00 - 15:00 lecure part 2b
* 16:00 - 17:00 exercises 2b [answers](part2_ex.html)

Schedule Fri:

* 9:00 - 10:00 lecture part 3a
* 10:00 - 10:15 coffee break
* 10:15 - 11:00 lecture part 3b
* 11:00 - 12:00 exercises 3 [answers](part3_ex.html)
* 12:00 - 13:00 lunch break
* 13:00 - 14:00 lecure part 4
* 14:00 - 15:00 exercises 4 [answers](part4_ex.html)

Further:

* please raise hands or speak up whenever something comes up
* please also share questions you run into in your actual research, preferably with (example) data and R code

### Syllabus

* **part 1**: the new spatial stack in R:
  simple features, DE-9IM, vector and raster data, data cubes;
  the spatial statistics data types
* **part 2** geometry operations, raster-vector, vector-raster:
  geometry measures, predicates, and transformers;
  spherical geometry;
  raster-vector: polygonizing, extracting;
  vector-raster: rasterize, interpolate, density;
  up- and down-scaling: aggregation, sampling, area-weighted interpolation, dasymetric mapping;
* **part 3**. inference: spatial correlation, fitting models:
  spatial correlation for point patterns, geostatistical data, and lattice data;
  fitting regression models under spatial correlation
* **part 4** prediction and simulation;
  point patterns: fitting densities, simulating point patterns;
  geostatistics: kriging interpolation, conditional simulation

### Resources

* [Spatial Data Science: With applications in R](https://r-spatial.org/book/), by Pebesma and Bivand 2023 (open online, hard copy from [CRC](https://www.routledge.com/Spatial-Data-Science-With-Applications-in-R/Pebesma-Bivand/p/book/9781138311183))
* Vignettes of [sf](https://r-spatial.github.io/sf/): tab "Articles"
* Vignettes of [stars](https://r-spatial.github.io/stars/): tab "Articles"

All these material were written using [quarto](https://quarto.org/) or R-markdown

Reading as preparation: students may want to read from this book chapters: 3-7, 10-12, 14-16

## Why R for spatial statistics / geospatial data science?

(skip)

* R is old! Think of the advantages!
* R is as good as any data science language, but is more in focus with the statistical community
* Most researchers in spatial statistics who share code have used or use R
* R has a strong ecosystem of users and developers, who communicate and collaborate (and compete, mostly in a good way)
* R spatial packages have gone full cycle: 
    * the first generation has been deprecated (`rgdal`, `rgeos`, `maptools`),
    * then removed from CRAN, and 
    * superseded by modern versions (`sf` and `stars` replaced `sp`, `terra` replaced `raster`)
* R is a data science language that allows you [to work reproducibly](https://eartharxiv.org/repository/view/3216/)
* Because we have [CRAN](https://cran.r-project.org/) and CRAN Taskviews: [Spatial](https://cran.r-project.org/web/views/Spatial.html), [SpatioTemporal](https://cran.r-project.org/web/views/SpatioTemporal.html), [Tracking](https://cran.r-project.org/web/views/Tracking.html)

### Reproducing or recreating the current course

(skip)

* Go to [https://github.com/edzer/ethz/](https://github.com/edzer/ethz/)
* Go to "Code", then "copy URL to clipboard"
* Clone this repo to your hard drive
* Start RStudio by double clickign the `ethz.Rproj` file in the source directory
* Reproduce these course materials by [installing quarto](https://quarto.org/docs/get-started/) and 
    * in RStudio: run `build` - `render book`, or
    * on the command line: run `quarto render` in the course directory 
* Run individual code sections in RStudio, and modify them!

## Exercises

1. Install the `spDataLarge` package (see instructions above)
2. Copy (or clone) the course material from GitHub to your local machine
3. Open it in RStudio
4. Open the `part1.qmd` file. Try to identify a code chunk.
5. Run the first code chunk.
6. Skip to the last code chunk; run all code chunks above it (by a single click), and then run this last code chunk.
7. Render the entire course "book", view the result by opening `_book/index.html` in a web browser (from Rstudio)
